The Impact of Alternation

### 2017-05-10 draft.tex
ok

  For the second point, we prove the following invariant: each time
  Algorithm \ref{alg:predabs} reaches line \ref{ln:while}, the set $W$
  of nodes in $\worklist$ contains all the frontier nodes in the ART
  $\tuple{N\cup W,E,\Lambda,\lhd}$ which are not covered by some node
  in $N$, namely that:
  \begin{equation}\label{inv:predabs}
  W = \set{n \mid \forall m \in N~ \forall a \in \Sigma ~.~ (n,a,m)
    \not\in E ~\wedge~ (n,m) \not\in \lhd}
  \end{equation}
  Initially, this is the case because $W = \set{\rootNode}$ and $N = E
  = \lhd = \emptyset$. If the invariant holds previously, at line
  \ref{ln:while}, it will hold again after line \ref{ln:refine} is
  executed, because, when the subtree rooted at the pivot $p$ is
  removed, $p$ becomes a member of the set of uncovered frontier
  nodes, and is added to $W$ at line \ref{ln:refine}. The last case
  considers that the invariant holds at line \ref{ln:while} and the
  control follows the else branch at line \ref{ln:else}. In this case,
  the newly created frontier node $s$ is added to $W$ only if it is
  not covered by an existing node in $N$ (line \ref{ln:covered}).

@@

  \EndWhile  
\end{algorithmic}

@@

  For the second point, we prove the following invariant: each time
  Algorithm \ref{alg:impact} reaches the line \ref{ln:impact-while},
  the set of nodes $W$ in $\worklist$ contains exactly those nodes in
  the ART $\tuple{N \cup W,E,\Lambda,R,T,\lhd}$ which are
  both\begin{inparaenum}[(i)]
  \item uncovered, and such that
  \item for some $a \in \Sigma$, the edge $(n,a,s)$ is missing from $E$:
  \end{inparaenum}
  \begin{equation}\label{inv:impact}
  W = \set{n \mid n \text{ is uncovered and } \exists a \in \Sigma~
    \forall s \in N ~.~ (n,a,s) \not\in E}
  \end{equation}
  Initially, $W = \set{\rootNode}$ and $N=E=\lhd=\emptyset$, thus the
  invariant holds trivially. Suppose that the invariant holds at some
  point, when reaching line \ref{ln:impact-while} and some node $n$
  was removed from $W$ and inserted into $N$. We distinguish two
  cases, either: \begin{itemize}
  \item $n$ is covered at line \ref{ln:close-worklist}, in which case
    $W$ remains unchanged, or
  %
  \item $n$ is not covered, in which case for each $a \in \Sigma$ a
    new node $s$ is inserted into $W$ (line \ref{ln:expand-end}) and
    the edge $(n,a,s)$ is inserted into $E$ (line
    \ref{ln:edge-insert}). None of these new nodes are covered and
    none of their successors inserted into $E$.
  \end{itemize}
  Then the condition (\ref{inv:impact}) holds next time the line
  \ref{ln:impact-while} is reached, thus proving the invariant.  \qed}

>>>>

  For the second point, we prove the following invariant: each time
  Algorithm \ref{alg:predabs} reaches line \ref{ln:while}, the set $W$
  of nodes in $\worklist$ contains all the frontier nodes in the ART
  $\tuple{N\cup W,E,\Lambda,\lhd}$ which are not covered by some node
  in $N$, namely that:
  \begin{equation}\label{inv:predabs}
  W = \set{n \not\in N \mid \forall m \in N~ \forall a \in \Sigma ~.~ (n,a,m)
    \not\in E ~\wedge~ (n,m) \not\in \lhd}
  \end{equation}
  Initially, this is the case because $W = \set{\rootNode}$ and $N = E
  = \lhd = \emptyset$. If the invariant holds previously, at line
  \ref{ln:while}, it will hold again after line \ref{ln:refine} is
  executed, because, when the subtree rooted at the pivot $p$ is
  removed, $p$ becomes a member of the set of uncovered frontier
  nodes, and is added to $W$ at line \ref{ln:refine}. The last case
  considers that the invariant holds at line \ref{ln:while} and the
  control follows the else branch at line \ref{ln:else}. In this case,
  the newly created frontier node $s$ is added to $W$ only if it is
  not covered by an existing node in $N$ (line \ref{ln:covered}).

@@

  \State {\bf return} $\true$
\end{algorithmic}

@@

  For the second point, we prove the following invariant: each time
  Algorithm \ref{alg:impact} reaches the line \ref{ln:impact-while},
  the set of nodes $W$ in $\worklist$ contains exactly those nodes in
  the ART $\tuple{N \cup W,E,\Lambda,R,T,\lhd}$ which are
  both\begin{inparaenum}[(i)]
  \item uncovered, and such that
  \item for some $a \in \Sigma$, the edge $(n,a,s)$ is missing from $E$:
  \end{inparaenum}
  \begin{equation}\label{inv:impact}
  W = \set{n \not\in N \mid n \text{ uncovered, } \exists a \in
    \Sigma~ \forall s \in N ~.~ (n,a,s) \not\in E}
  \end{equation}
  Initially, $W = \set{\rootNode}$ and $N=E=\lhd=\emptyset$, thus the
  invariant holds trivially. Suppose that (\ref{inv:impact}) holds at
  when reaching line \ref{ln:impact-while} and some node $n$ was
  removed from $W$ and inserted into $N$. We distinguish two cases,
  either: \begin{itemize}
  \item $n$ is covered at line \ref{ln:close-worklist}, in which case
    $W$ remains unchanged, or
  %
  \item $n$ is not covered, in which case for each $a \in \Sigma$ a
    new node $s$ is inserted into $W$ (line \ref{ln:expand-end}) and
    the edge $(n,a,s)$ is inserted into $E$ (line
    \ref{ln:edge-insert}). None of these new nodes are covered and
    none of their successors inserted into $E$.
  \end{itemize}
  Then the condition (\ref{inv:impact}) holds next time line
  \ref{ln:impact-while} is reached, thus it is invariant. If Algorithm
  \ref{alg:impact} returns $\true$ then $W = \emptyset$ and, by
  (\ref{inv:impact}), for each node in $n \in N$ one of the following
  hold: \begin{compactitem}
  \item $n$ is covered, or
  \item for each $a \in \Sigma$ there exists $s \in N$ such that
    $(n,a,s) \in E$. 
  \end{compactitem}
  We prove that, in this case, $\bigvee_{n \in N} \Lambda(n)$ defines
  a safety invariant and conclude that $L(\A) = \emptyset$, by Lemma
  \ref{lemma:safety-invariant}. To show that $\bigvee_{n \in N}
  \Lambda(n)$ defines an invariant, let $u \in \Sigma^*$ be an
  arbitrary sequence. \qed}
